\chapter{Fitted Q-Iteration with Deep State Features}
\label{ch3_setup}
\thispagestyle{empty}

\vspace{0.5cm}

As central argument of this thesis, we propose a DRL method which 
combines the feature extraction capabilities of deep CNNs with the quick 
and powerful batch RL approach of FQI. 
Given a high-dimensional state space of pixels representing sequences of 
greyscale frames from an Atari game, we use a deep convolutional autoencoder 
to map the original state space to a compressed \textit{feature space} which 
accounts for the states and its nominal state dynamics (i.e. those changes not
directly influenced by the agent). We reduce the representation further by 
applying a feature selection algorithm to the extracted state features and this 
final compressed representation is then used to run a \textit{tree-based} 
version of FQI. We repeat this procedure iteratively resulting in a 
\textit{semi-batch} algorithm. 

In this chapter we give a formal description of the method and its core 
components. Technical details of implementation will be discussed in the next
chapter.

\section{Motivation}
The state-of-the-art DRL methods listed in the previous chapter are able to 
outperform classic RL algorithms in a wide variety of problems, and in some 
cases are the only possible way of dealing with high-dimensional control 
settings like the Atari games. 
However, the approaches cited above tend to be grossly 
\textit{sample-inefficient}, requiring tens of millions of samples collected
on-line to reach optimal performance. Several publications successfully deal 
with this aspect, but nonetheless leave room for improvement (lowering at most
by one order of magnitude the number of samples required).
The method introduced by Lange and Riedmiller (2010) \cite{lange2010deep} is 
similar to ours but their dense architecture predates the more modern 
convolutional approaches in image processing and is less suited for complex
tasks than our AE.

The method that we propose tries to improve both aspects of information content
of the compressed feature space and sample efficiency. We extract general 
features from the environments and try to reach better or equivalent performance
in up to two orders of magnitude less samples than DQN on Atari games.

\section{Algorithm Description}
The general structure of this algorithm is typical of DRL settings: we use a 
deep ANN to extract a representation of an environment, and use that 
representation to control an agent with standard RL algorithms. We also add an 
additional step after the deep feature extraction to further reduce the 
representation down to the essential bits of information required to solve the 
problem by using the \textit{Recursive Feature Selection} (RFS) algorithm 
\cite{castelletti2011tree}.

We focus exclusively on environments with a discrete and mono-dimensional 
action space $A$, where actions are assigned a unique integer identifier 
starting from $0$ with no particular order. 
We also assume to be operating in a three dimensional state space 
($C \times W \times H$\footnote{Channels times Width times Height}) for 
consistency with the experimental setting on which we tested the algorithm (with
CNNs and pixel-level state spaces), although in general the algorithm requires 
no such assumption and could be easily adapted to higher or lower dimensional 
settings. 

The algorithm uses a modular architecture which results in three separate stages
for the training phase, and combines the modules in an end-to-end fashion during
the control phase. The main components of the algorithm are:
%
\begin{enumerate}
    \item a deep convolutional autoencoder which we use to extract a 
    representation of the environment;
    the purpose of the AE is to map the original, pixel-level state space $S$ of
    the environment into a strongly compressed feature space ${\tilde{S}}$ which
    contains information of both the state space and part of the transition 
    model of the environment;
    \item the \textit{Recursive Feature Selection} (RFS) technique to further 
    reduce the state representation $\tilde{S}$ and keep only the truly 
    informative features extracted by the AE, effectively mapping the extracted 
    state-space to a subspace $\hat{S}$.
    \item the \textit{tree-based} FQI learning algorithm which produces an 
    estimator for the action-value function, with $\hat{S}$ as domain. 
\end{enumerate}
%

At each training step, the modules are sequentially trained to produce the 
following transformations:
\begin{itemize}
    \item $ENC: S \rightarrow \tilde{S}$, from the pixel representation to a 
    compressed feature space;
    \item $RFS: \tilde{S} \rightarrow \hat{S}$, from the compressed feature space
    to a minimal subspace with the most informative features;
    \item $\hat{Q}: \hat{S} \times A \rightarrow \mathbb{R}$, an approximation
    of the optimal action-value function on $\hat{S}$.
\end{itemize}
After the training phase, we simply combine the three functions to obtain the 
real action-value function $Q: S \times A \rightarrow \mathbb{R}$ as follows: 
%
\begin{IEEEeqnarray}{rCl}
    %
    Q(s, a) = \hat{Q}(RFS(ENC(s)), a) \label{eq:final_output}
    %
\end{IEEEeqnarray}
%

A full training step of the algorithm takes as input a training set 
$\mathcal{TS}$ of four-tuples $(s \in S, a \in A, r \in \mathbb{R}, s' \in S)$
and produces a new approximation of the action-value function.
We define a greedy policy $\pi$ based on the current approximation of $Q$ as:
%
\begin{IEEEeqnarray}{rCl}
    %
    \pi(s) = \underset{a}{\arg\max} Q(s, a)
    %
\end{IEEEeqnarray}
%
and we use an $\varepsilon$-greedy policy $\pi_\varepsilon$ based on $\pi$ to 
collect $\mathcal{TS}$ at each training step, in a \textit{semi-batch} way.
% TODO: talk about how you initialize the policy to random, maybe split in more paragraphs
After each step, we decrease the randomicity of $\pi_\varepsilon$ and allow the
agent to exploit the learned $Q$ to collect the new training set. This 
results in a sufficiently high exploration at the beginning of the procedure, 
but increasingly exploits the learned knowledge to improve the quality of the 
collected samples after each training step, as the agents sees better and rarer
states. 
A similar approach, called \textit{$\varepsilon$-annealing}, was used by Mnih et
al.\ (2015) \cite{mnih2015human} for the online updates in DQN.
Each training step is followed by an evaluation of the greedy policy $\pi$ to 
determine the quality of the learned $Q$ and eventually stop the procedure
when the performance is deemed satisfactory. 

A general description of the process is reported in Algorithm \ref{alg:FQI-DSDF},
and details on each training phase are given in the following sections. 
%
\begin{algorithm}[h]
    \caption{Fitted Q-Iterations with Deep State Features}
    \label{alg:FQI-DSDF}
    \begin{algorithmic}
	\STATE Initialize the encoder $ENC: S \rightarrow \tilde{S}$ arbitrarily;
	\STATE Initialize the decoder $DEC: \tilde{S} \rightarrow S$ arbitrarily;
	\STATE Initialize $Q$ arbitrarily;
	\STATE Define $\pi(s) = \underset{a}{\arg\max} Q(s, a)$;
	\STATE Initialize an $\varepsilon$-greedy policy $\pi_\varepsilon$ based on $\pi$ with $\varepsilon = 1$;
	\REPEAT 
	    \STATE Collect a set $\mathcal{TS}$ of four-tuples $(s \in S, a \in A, r \in \mathbb{R}, s' \in S)$ using $\pi_\varepsilon$;
	    \STATE Train the composition $DEC \circ ENC: S \rightarrow S$ using the first column of $\mathcal{TS}$ as input and target;
	    \STATE Build a set $\mathcal{TS}_F$ of four-tuples $(f \in \tilde{S}, a \in A, r \in \mathbb{R}, f' \in \tilde{S})$ by applying the encoder to the first and last column of $\mathcal{TS}$ s.t. $f = ENC(s)$;
	    \STATE Call the RFS feature selection algorithm on $\mathcal{TS}_F$ to obtain a space reduction $FS: \tilde{S} \rightarrow \hat{S}$;
	    \STATE Build a set $\mathcal{TS}_{\hat{F}}$ of four-tuples $(\hat{f} \in \hat{S}, a \in A, r \in \mathbb{R}, \hat{f'} \in \hat{S})$ by applying $RFS$ to the first and last column of $\mathcal{TS}_F$ s.t. $\hat{f} = RFS(f)$;
	    \STATE Call FQI on $\mathcal{TS}_{\hat{F}}$ to produce $\hat{Q}: \hat{S} \times A \rightarrow \mathbb{R}$;
	    \STATE Combine $\hat{Q}$, $RFS$ and $ENC$ to produce $Q: S \times A \rightarrow \mathbb{R}$:
		\[
		    Q(s, a) = \hat{Q}(RFS(ENC(s)), a)
		\]
	    \STATE Reduce $\varepsilon$ according to some criterion;
	\UNTIL{stopping condition is met;}
	\STATE Return $Q$
    \end{algorithmic}
\end{algorithm}
%

\section{Extraction of State Features}
The AE used in our approach consists of two main components, namely an 
\textit{encoder} and a \textit{decoder} (cf.\ Section \ref{s:AE}). To the end of 
explicitly representing the encoding purpose of the AE, we keep a separate 
notation of the two modules; we therefore refer to two different CNNs, namely 
$ENC: S \rightarrow \tilde{S}$ that maps the original state space to the 
compressed representation $\tilde{S}$, and $DEC: \tilde{S} \rightarrow S$ which
performs the inverse transformation. The full AE is the composition of the two 
networks $AE: DEC \circ ENC: S \rightarrow S$. Note that the composition is 
differentiable end-to-end, and basically consists in \textit{plugging} the last
layer of the encoder as input to the decoder. 

We train the AE ...

\section{Tree-based Recursive Feature Selection}
% TODO

\section{Fitted Q-Iteration}
The last component of our training pipeline is the Fitted Q-Iteration algorithm 
(cf. \ref{s:FQI}) run on the training set $\mathcal{TS}_{\hat{F}}$ of 
four-tuples $(\hat{f} \in \hat{S}, a \in A, r \in \mathbb{R}, \hat{f'} \in \hat{S})$, 
obtained by applying RFS to the first and last column of 
$\mathcal{TS}_{\tilde{F}}$. 
We train the model to output a multi-dimensional estimate on $\mathbb{R}^{|A|}$,
with one value for each action that the agent can take in the environment, and 
we select the single output value on $\mathbb{R}$ using the action identifier as
index (e.g.\ $Q(s, 0)$ will return the first dimension of the model's output). 
The final output of the training procedure is an estimator 
$\hat{Q}^\pi: \hat{S} \times A \rightarrow \mathbb{R}$ which maps the compressed
and reduced state-action space to a real value. 

% TODO: image of the full model AE + RFS + FQI
 
The final step of our algorithm consists in composing the outputs of the three 
modules in a single, end-to-end model which can operate directly on the original
feature space $S$ as defined in Equation \eqref{eq:final_output}.
